---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  fig.width=7, 
  fig.height=6
)
```

```{r setup}
library(taldi)
library(caracas)
```

# Summary

1. `infer_ast()`: infer abstract syntex tree
2. `make_dag()`: construct directed acyclic graph
3. (`collect_leaves()`: Collect symbol leaves)
4. Compute
  * `forward_computation()`
  * `reverse_ad()`

# Infering abstract syntex tree

```{r}
e <- quote(cos(2*x1 + x2) + x2^2)
symlist <- infer_ast(e)
str(symlist, 2)
```

Construct directed acyclic graph (DAG):

```{r}
dag <- make_dag(symlist)

```

Plot DAG:

```{r}
ggdag(dag)
```


Collect symbol leaves (such that each symbol only occurs once):

```{r}
dag2 <- collect_leaves(dag)
ggdag(dag2)
```


# Evaluating function in point

Forward computation:

```{r}
e <- quote(cos(2*x1 + x2) + x2^2)
vals <- list(x1 = 1, x2 = 2)
ast <- infer_ast(e)
dag <- make_dag(ast)
dag <- collect_leaves(dag)
dag <- forward_computation(dag, values = vals)
```

The value in the node is in the second line in each node:

```{r}
ggdag(dag)
```

The numeric value can be extracted as follows:

```{r}
root <- get_root(dag)
get_ad_values(dag, root)
eval(e, vals)
```




# Finding derivatives

## Symbolically

```{r}
e <- quote(cos(2*x1 + x2) + x2^2)
vals <- list(x1 = 1, x2 = 2)
```

Derivatives can be found symbolically using `caracas` (`R` code hidden here, please refer to the vignette source for details):

```{r, echo = FALSE}
f <- as_sym(deparse1(e))
#f
dfdx1 <- der(f, "x1")
dfdx2 <- der(f, "x2")
```

\begin{align}
  f(x_1, x_2) &= `r tex(f)` \\
  \frac{\partial f}{\partial x_1} &= `r tex(dfdx1)` \\
  \frac{\partial f}{\partial x_2} &= `r tex(dfdx2)`
\end{align}

And evaluated in the point $(x_1, x_2) = (`r vals["x1"]`, `r vals["x2"]`)$:

```{r, echo = FALSE}
#val <- eval_to_symbol("UnevaluatedExpr(pi/2)")
#val1 <- eval_to_symbol("UnevaluatedExpr(1)")
#val2 <- eval_to_symbol("UnevaluatedExpr(2)")
val1 <- eval_to_symbol(paste0("UnevaluatedExpr(", vals[1L], ")"))
val2 <- eval_to_symbol(paste0("UnevaluatedExpr(", vals[2L], ")"))

f_val_uneval <- subs_lst(f, list("x1" = val1, "x2" = val2))
#f_val_uneval
f_val <- doit(f_val_uneval)
#f_val

dfdx1_val_uneval <- subs_lst(dfdx1, list("x1" = val1, "x2" = val2))
#dfdx1_val_uneval
dfdx1_val <- doit(dfdx1_val_uneval)
#dfdx1_val

dfdx2_val_uneval <- subs_lst(dfdx2, list("x1" = val1, "x2" = val2))
#dfdx2_val_uneval
dfdx2_val <- doit(dfdx2_val_uneval)
#dfdx2_val
```

\begin{align}
  f\left (`r tex(val1)`, `r tex(val2)` \right ) 
    &= `r tex(f_val_uneval)` 
     = `r tex(f_val)` 
     \approx `r N(f_val, 5)`\\
  \frac{\partial f}{\partial x_1} \Bigg \vert_{x_1 = x_2 = `r tex(val1)`}
    &= `r tex(dfdx1_val_uneval)` 
     = `r tex(dfdx1_val)` 
     \approx `r N(dfdx1_val, 5)` \\
  \frac{\partial f}{\partial x_2} \Bigg \vert_{x_1 = x_2 = `r tex(val2)`}
    &= `r tex(dfdx2_val_uneval)` 
     = `r tex(dfdx2_val)` 
     \approx `r N(dfdx2_val, 5)` 
\end{align}

## Reverse algorithmic differentiation (AD)

Using reverse algorithmic differentiation (AD):

```{r}
ast <- infer_ast(e)
dag <- make_dag(ast)
dag <- collect_leaves(dag)
dag <- reverse_ad(dag, values = vals)
ggdag(dag)
```


# Advanced 

## Find leaves

```{r}
is_leaf <- sapply(V(dag), function(x) {
  length(neighbors(dag, x, mode = "out")) == 0L
})
V(dag)[is_leaf]
leaves_labels <- vertex_attr(dag, "label", V(dag)[is_leaf])
leaves_types <- vertex_attr(dag, "type", V(dag)[is_leaf])
cbind(leaves_labels, leaves_types)
```



## Preparing

* `init_graph()`
* `bind_literals()`
* `bind_symbols()`

```{r}
ast <- infer_ast("cos(2*x1 + x2) + x2^2")
dag <- make_dag(ast)
dag <- collect_leaves(dag)
```

Init all node values to `NA`:

```{r}
dag2 <- init_graph(dag)
ggdag(dag2)
```

Bind literals:

```{r}
dag3 <- bind_literals(dag2)
ggdag(dag3)
```

Bind symbols:

```{r}
dag4 <- bind_symbols(dag2, values = list(x1 = 1, x2 = 2))
ggdag(dag4)
```

Bind both literals and symbols:

```{r}
dag5 <- bind_literals(dag2)
dag5 <- bind_symbols(dag5, values = list(x1 = 1, x2 = 2))
ggdag(dag5)
```



## Plot using `ggraph` and `tidygraph`


```{r}
library(tidygraph)
library(ggraph)
```

```{r}
ast <- infer_ast("cos(2*x1 + x2) + x2^2")
dag <- make_dag(ast)
dag <- collect_leaves(dag)

tg <- as_tbl_graph(dag)
l <- create_layout(tg, layout = "sugiyama")

p <- ggraph(l) + 
  geom_edge_link(arrow = arrow(length = unit(0.25, 'cm')), end_cap = circle(0.85, 'cm')) + 
  geom_node_circle(aes(r = 0.3, fill = type)) + 
  geom_node_text(aes(label = label)) + 
  labs(fill = NULL) + 
  theme_bw() + 
  theme_graph()

print(p)  
```

This is essentially how `ggdag()` is implemented.


