---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(taldi)
library(tidygraph)
library(ggraph)
library(caracas)
```

# Summary

1. `infer_ast()`: infer abstract syntex tree
2. `make_dag()`: construct directed acyclic graph
3. (`collect_leaves()`: Collect symbol leaves)

# Infering abstract syntex tree

```{r}
e <- quote(cos(2*x1 + x2) + x2^2)
symlist <- infer_ast(e)
str(symlist, 2)
```

# Construct directed acyclic graph

```{r, fig.width=6, fig.height=4}
dag <- make_dag(symlist)
plot(dag)
```

## Find leaves

```{r}
is_leaf <- sapply(V(dag), function(x) {
  length(neighbors(dag, x, mode = "out")) == 0L
})
V(dag)[is_leaf]
leaves_labels <- vertex_attr(dag, "label",  V(dag)[is_leaf])
leaves_types <- vertex_attr(dag, "type",  V(dag)[is_leaf])
cbind(leaves_labels, leaves_types)
```


# Plot using `ggraph` and `tidygraph`

```{r, fig.width=6, fig.height=4}
tg <- as_tbl_graph(dag)
l <- create_layout(tg, layout = "sugiyama")

p <- ggraph(l) + 
  geom_edge_link(arrow = arrow(length = unit(0.5, 'cm')), end_cap = circle(0.5, 'cm')) + 
  geom_node_circle(aes(r = 0.1, fill = type)) + 
  geom_node_text(aes(label = label)) + 
  labs(fill = NULL) + 
  theme_bw() + 
  theme_graph()

print(p)  
```

Or simply use the provided function, `ggdag()`:

```{r, fig.width=6, fig.height=4}
ggdag(dag)
```


# Collect symbol leaves

```{r}
dag2 <- collect_leaves(dag)
ggdag(dag2)
```


# Complete example

```{r, fig.width=6, fig.height=4}
ast <- infer_ast("cos(2*x1 + x2) + x2^2")
dag <- make_dag(ast)
dag <- collect_leaves(dag)
dag2 <- reverse_ad(dag, values = list(x1 = 1, x2 = 2))
ggdag(dag2)
```

# Low-level functions

## Preparing

* [init_graph()]
* [bind_literals()]
* [bind_symbols()]

```{r}
ast <- infer_ast("cos(2*x1 + x2) + x2^2")
dag <- make_dag(ast)
dag <- collect_leaves(dag)
```

```{r, fig.width=6, fig.height=4}
ggdag(dag)
```


```{r}
dag2 <- init_graph(dag)
```

```{r, fig.width=6, fig.height=4}
dag3 <- bind_literals(dag2)
ggdag(dag3)
```


```{r, fig.width=6, fig.height=4}
dag4 <- bind_symbols(dag2, values = list(x1 = 1, x2 = 2))
ggdag(dag4)
```


```{r}
dag5 <- bind_literals(dag2)
dag5 <- bind_symbols(dag5, values = list(x1 = 1, x2 = 2))
```

```{r, fig.width=6, fig.height=4}
ggdag(dag5)
```


```{r, fig.width=6, fig.height=4}
dag5 <- forward_computation(dag5, values = list(x1 = 1, x2 = 2))
ggdag(dag5)
```


## Forward computation

```{r}
e <- quote(cos(2*x1 + x2) + x2^2)
vals <- list(x1 = 1, x2 = 2)
eval(e, vals)
ast <- infer_ast(e)
dag <- make_dag(ast)
dag <- collect_leaves(dag)
dag <- forward_computation(dag, values = vals)
```

```{r, fig.width=6, fig.height=4}
ggdag(dag)
```


# Find derivative

```{r}
f <- as_sym(deparse1(e))
f
dfdx1 <- der(f, "x1")
dfdx2 <- der(f, "x2")
```

\begin{align}
  f(x_1, x_2) &= `r tex(f)` \\
  \frac{\partial f}{\partial x_1} &= `r tex(dfdx1)` \\
  \frac{\partial f}{\partial x_2} &= `r tex(dfdx2)`
\end{align}

```{r}
val <- eval_to_symbol("UnevaluatedExpr(pi/2)")

f_val_uneval <- subs_lst(f, list("x1" = val, "x2" = val))
f_val_uneval
f_val <- doit(f_val_uneval)
f_val

dfdx1_val_uneval <- subs_lst(dfdx1, list("x1" = val, "x2" = val))
dfdx1_val_uneval
dfdx1_val <- doit(dfdx1_val_uneval)
dfdx1_val

dfdx2_val_uneval <- subs_lst(dfdx2, list("x1" = val, "x2" = val))
dfdx2_val_uneval
dfdx2_val <- doit(dfdx2_val_uneval)
dfdx2_val
```

\begin{align}
  f\left (`r tex(val)`, `r tex(val)` \right ) 
    &= `r tex(f_val_uneval)` 
     = `r tex(f_val)`  \\
  \frac{\partial f}{\partial x_1} \Bigg \vert_{x_1 = x_2 = `r tex(val)`}
    &= `r tex(dfdx1_val_uneval)` 
     = `r tex(dfdx1_val)`  \\
  \frac{\partial f}{\partial x_2} \Bigg \vert_{x_1 = x_2 = `r tex(val)`}
    &= `r tex(dfdx2_val_uneval)` 
     = `r tex(dfdx2_val)`
\end{align}

